---
title: "1st Assignment"
author: "Somtoo Henry"
format: pdf
editor: visual
---

**Q1**

Install the packages tidyverse, haven and dplyr if you do not have it

```{r}
install.packages("tidyverse")
install.packages("haven")
install.packages("dplyr")
```

Load the packages tidyverse, haven and dplyr

```{r}
library(tidyverse)
library(haven)
library(dplyr)
```

Import the data avengers

```{r}
avengers <- read.csv("avengers.csv") #import the dataset and assign to object
```

View the data avengers

```{r}
head(avengers)   #display the first 6 rows of the dataset
view(avengers)   #view the dataset avengers
```

Find the total number of missing values in the data avengers

```{r}
sum(is.na(avengers)) #sum the total missing values
```

Find the number of missing values per column

```{r}
colSums(is.na(avengers))  #sum the missing values per column
```

**Q2**

Clean the data to remove missing values and retain complete cases, and assign to a new object

```{r}
Cleanset <- avengers %>% #used to transfer or connect the dataset to the function
  na.omit() #clean the dataset by removing missing values

head(Cleanset)#Display the first 6 rows of the cleaned the dataset 
view(Cleanset)#View the dataset
```

Subset the data to include CombatEffectiveness and assign it to the object Cleandata

```{r}
Cleandata <- avengers %>%  #transfer the dataset to the function
  na.omit() %>%   # Remove the missing values and tranfer the dataset to mutate 
  mutate(CombatEffectiveness=(agility+speed+strength+willpower)) #Adds a new variable to the dataset
 
View(Cleandata)#view the dataset  
head(Cleandata)#Display the first 6 rows of the dataset
```

**Q3**

Create a new dataset of avengers that did not have superpowers and died to export in csv and SPSS.

```{r}
Datacopy<- Cleandata%>%   #Transfer the dataset to filter
  filter(superpower=="no"& died=="yes") #filter avengers with no superpowers and those that died

view(Datacopy)   #View the dataset
head(Datacopy)   #Display the first 6 rows of the dataset
```

Export the new dataset of avengers that did not have superpowers and died in csv format.

```{r}
write.csv(Datacopy,"Datacopy.csv",row.names = FALSE,na="") #Exports dataset in csv format
```

Export the new dataset of avengers that did not have superpowers and died in SPSS format.

```{r}
write_sav(Datacopy,"Datacopy.sav")  #Export dataset in SPSS format
```

Summarise the new dataset "Datacopy" to report the mean, SD, and range values for the overall sample

```{r}
Generalsum <-Datacopy %>% #transfer dataset to summarise function
  summarise (
    N = n(), #The total number of avengers in Datacopy
    
    # Find the average of combat effectiveness
    avg_combat = mean(CombatEffectiveness, na.rm = TRUE), 
    # Find the SD of combat effectiveness
    sd_combat = sd(CombatEffectiveness, na.rm = TRUE),    
    # Find the Minimum range of combat effectiveness
    min_combat = min(CombatEffectiveness, na.rm = TRUE),  
    # Find the Maximum range combat effectiveness
    max_combat = max(CombatEffectiveness, na.rm = TRUE), 
    
    # Find the average of kills
    avg_kills = mean(kills, na.rm = TRUE),
    #Find the SD of kills
    sd_kills = sd(kills, na.rm = TRUE),
    # Find the minimum range of kills
    min_kills = min(kills, na.rm = TRUE),
    #Find the maximum range of kills
    max_kills = max(kills, na.rm = TRUE),                
    
    # Find the average of injuries
    avg_injuries = mean(injuries, na.rm = TRUE),
    # Find the SD of injuries
    sd_injuries = sd(injuries, na.rm = TRUE),
    #Find the Minimum range of injuries
    min_injuries = min(injuries, na.rm = TRUE),
    # Find the Maximum range of injuries
    max_injuries = max(injuries, na.rm = TRUE))           
    
print(Generalsum) #Display the new data set assigned to object "Generalsum"
 
```

Summarise the new dataset "Datacopy" based on battlefield location (i.e., north vs. south)

```{r}
Sumlocation <- Datacopy %>% #transfer the dataset to group by function
  group_by(north_south) %>% # Group the dataset "Datacopy" by battlefield location
  summarise(                #i.e north and south
    N = n(),# Total number per group
   
    # Find the average of combat effectiveness
    avg_combat = mean(CombatEffectiveness, na.rm = TRUE), 
    # Find the SD of combat effectiveness
    sd_combat = sd(CombatEffectiveness, na.rm = TRUE),    
    # Find the Minimum range of combat effectiveness
    min_combat = min(CombatEffectiveness, na.rm = TRUE),  
    # Find the Maximum range combat effectiveness
    max_combat = max(CombatEffectiveness, na.rm = TRUE), 
    
    # Find the average of kills
    avg_kills = mean(kills, na.rm = TRUE),
    #Find the SD of kills
    sd_kills = sd(kills, na.rm = TRUE),
    # Find the minimum range of kills
    min_kills = min(kills, na.rm = TRUE),
    #Find the maximum range of kills
    max_kills = max(kills, na.rm = TRUE),                
    
    # Find the average of injuries
    avg_injuries = mean(injuries, na.rm = TRUE),
    # Find the SD of injuries
    sd_injuries = sd(injuries, na.rm = TRUE),
    #Find the Minimum range of injuries
    min_injuries = min(injuries, na.rm = TRUE),
    # Find the Maximum range of injuries
    max_injuries = max(injuries, na.rm = TRUE))           
     
print(Sumlocation)#Display the data assigned to object sumlocation
```

Export the summary stats for overall data and based on location in csv

```{r}
write.csv(Generalsum,"Generalsum.csv",row.names = FALSE,na="") #Export generalsum in csv format
write.csv(Sumlocation,"Sumlocation.csv",row.names = FALSE,na="") #Export sumlocation in csv format
```

**Q4**

Find the battlefield that is most effective in combat based on the battlefield location.

```{r}
#Select row with max avg combat from the dataset Sumlocation and assign to object
Highestcombat <- Sumlocation %>%
  filter(avg_combat == max(avg_combat))

print(Highestcombat)#Display the data assigned to object Highestcombat
```

Find the battlefield with the most injuries based on the battlefield location.

```{r}
#Select row with max avg injuries from the dataset Sumlocation and assign to object
Highestinjuries <- Sumlocation  %>%
  filter(avg_injuries == max(avg_injuries))

print(Highestinjuries)#Display the data assigned to object Highestinjuries

```

**Q5**

Find the most erroneous variable in the mean model

```{r}
#Summarise the mean,standard deviation and Coefficient of Variance for Combateffectiveness,kills and injuries and asssign to object.
Dataerror <- Datacopy %>% #Connect the dataset to the summarise function
  summarise(
    #The average of combat
    avg_combat = mean(CombatEffectiveness, na.rm = TRUE),
    #The of SD CombateEffectiveness
    sd_combat = sd(CombatEffectiveness, na.rm = TRUE),
    #Find the cv for CombatEffectiveness
    cv_combat = (sd_combat / avg_combat),
    
    #The average of kill
    avg_kills = mean(kills, na.rm = TRUE),
    #The SD of kills
    sd_kills = sd(kills, na.rm = TRUE),
    #Find the cv for kills
    cv_kills = (sd_kills / avg_kills),
    
    #The average of injuries
    avg_injuries = mean(injuries, na.rm = TRUE),
    #The SD of injuries
    sd_injuries = sd(injuries, na.rm = TRUE),
    #Find the CV for injuries
    cv_injuries = (sd_injuries / avg_injuries))

print(Dataerror)#Display the data assigned to object dataerror

#Export dataerror in csv format
write.csv(Dataerror,"Dataerror.csv",row.names = FALSE,na="")

#create a Vector of CV values from dataerror and assign to object
Datacv <- c(Combat = Dataerror$cv_combat,#Extract cv for combat from dataerror
  Kills = Dataerror$cv_kills,            #Extract cv for kills from dataerror
  Injuries = Dataerror$cv_injuries)      #Extract cv for injuries from dataerror

#find the variable with the most error
Datacv[which.max(Datacv)]
```

**Q8**

Carry out a power analysis with estimate target effect size

```{r}

#Install power package if you do not have it
install.packages(pwr)

#Load the package pwr
library(pwr)

# Run a Power analysis to determine required sample size 
Powertest <- pwr.t.test(
  n=NULL,                     #Observation in each group
  d = 0.5,                    # Cohens.d hypothetical Effect size (Medium)
  sig.level = 0.05,           # Type I probability
  power = 0.80,               # Desired power (1 minus Type II probability)
  type = "two.sample",        # Two samples t-test
  alternative = "two.sided")  # Two-sided t- test

# Display power analysis results
print(Powertest) #Display the result assigned to the object Powertest

```

**Q9**

Carry out an equivalence test to confirm enough power

```{r}
library(TOSTER)  #load the package TOSTER
#install the package "TOSTER" if you do not have it
install.packages("TOSTER")

# Calculate the required sample size for 90% power equivalence test
powerTOSTtwo(
  alpha             = 0.05,   #significance level
  statistical_power = 0.90,   # power for equivalence testing
  low_eqbound_d     = -0.5,   # Low equivalent bound (Cohen's d "medium") 
  high_eqbound_d    =  0.5)   # high equivalent bound(Cohen's d "medium")

# Calculate the achieved power
powerTOSTtwo(
  alpha             = 0.05,   # Type I error rate (significance level)
  N                 = 406,    # n per group (812/2)
  low_eqbound_d     = -0.5,   #low equivalent bound (Cohen's d "medium")
  high_eqbound_d    =  0.5)   #high equivalent bound(Cohen's d "medium")
```

**Q10**

Calculate effect-size with 95% confidence interval and give the effect a qualitative label using Cohen’s conventions.

```{r}
#Install package "effectsize" if you do not have it
install.packages("effectsize") #code to install package 

library(effectsize) #load package effectsize

# Effect size from t-statistic = 4.25

t_stat  <- 4.25  #t-statistic
n1      <- 406   #group 1
n2      <- 406   #group 2

#Calculate Cohen's d

CohensD <- t_stat * sqrt((n1 + n2) / (n1 * n2)) #calculate cohen d and assign result to the object
cat("Cohen's d =", round(CohensD, 3), "\n")

# Calculate  Cohen's d with 95% CI

CIresult <- t_to_d(
  t        = 4.25,        # observed t-statistic
  df_error = 812 - 2,     # degrees of freedom (N - 2)
  paired   = FALSE,       # independent samples t-test
  ci       = 0.95)        # 95% confidence interval)

print(CIresult) #Display CIresult

# Description of Cohen’s d qualitative label

cohen.ES(test = "t", size = "small")  #small effect
cohen.ES(test = "t", size = "medium") #medium effect
cohen.ES(test = "t", size = "large")  #large effect

```
